The Definite Extension Procedure for Large-Scale Marginal Satisfiability

### 2016-11-10 documents/DefiniteExtensions/dep.tex
Started work on slides for Brazil

% === Causal structure formatting for tikz ===
\input{../../figures/causal_structures/formatting}

\begin{document}
    \title{Inequalities Witnessing Quantum Incompatibility in The Triangle Scenario}
    \author{Thomas C. Fraser}
    \email{tcfraser@tcfraser.com}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada \\ University of Waterloo, Waterloo, Ontario, Canada}
    % \author{Elie Wolfe}
    % \email{ewolfe@perimeter@institute.ca}
    % \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada}
    \date{\today}
    \begin{abstract}
        Quantum correlations are often incompatible with a classical assumption of causal structure. This nonclassicality is often known as quantum nonlocality, and it is witnessed through the violation of causal compatability inequalities, such as Bell inequalities. Such inequalities were recently derived for the Triangle scenario [arXiv:1609.00672], begging the question: can these inequalities be violated by quantum correlations? Here we answer this affirmatively, and discuss specific Triangle scenario inequalities and quantum configurations which manifest nonclassical correlations. We also report the development of novel computational techniques suitable for large causal structures with numerous outcomes. Numerical optimzations reveal quantum resources potentially qualitatively different from those known previously.
    \end{abstract}
    \maketitle
    \tableofcontents

    \section{Introduction}
    \todo[TC]{Mention Bell's work} \\

    \todo[TC]{Violations of inequalities correspond to resources} \\

    \todo[TC]{Finding inequalities is hard} \\

    \todo[TC]{Causal inference is a great lens to answer these questions} \\

    \todo[TC]{Demonstrated the usefulness of the inflation technique} \\

    \todo[TC]{Definite Extension procedure} \\

    \todo[TC]{Optimizations against inequalities} \\

    \section{Compatibility, Contextuality and the Marginal Problem}
    \label{sec:comp_con_mp}

    \subsection{Preliminary Notions}
    \label{eq:comp_con_mp_notions}
    A reoccurring motif of these discussions will be a set of marginal probability distributions $\bc{\prob[V_1], \ldots, \prob[V_k]}$. In agreement with~\cite{Fritz_2011} we adopt the following definitions and notation.

    \begin{definition}
        \label{def:marginal_model}
        A set of random variables $\jointvar = \bc{v_1, \ldots, v_n}$ is referred to as the \term{joint random variables} when denoted with a $\jointvar$. A subset of joint variables $V \subset \jointvar$ is called \term{marginal context}. A collection of marginal contexts is called the \term{marginal scenario}\footnote{Rigorously speaking as defined by~\cite{Fritz_2011}, a marginal scenario forms an \textit{abstract simplicial complex} where it is required that $V' \in \mscenario$ whenever $V' \subseteq V \in \mscenario$. Throughout this document, we only consider the \textit{maximal marginal scenario}; restricting our focus to the largest contexts in the marginal scenario $\forall i,j \in 1, \ldots, k: V_i \not \subseteq V_j$.} and is denoted $\mscenario$\footnote{It is also assumed that the marginal scenario covers the joint variables $\jointvar = \bigcup_{i = 1}^{k} V_i$.}.
        \[ \mscenario \defined \bc{V_1, \ldots, V_k \mid V_i \subseteq \jointvar} \]
        Additionally, the set of probability distributions $\bc{\prob[V_1], \ldots, \prob[V_k]}$ over the marginal scenario is called a \term{marginal model} is compactly denoted $\prob^{\mscenario}$.
    \end{definition}
    A specific marginal model $\prob^{\mscenario}$ for a marginal scenario $\mscenario$ abstractly represents a set of measurements or observations made on a physical system. Associated with the physical system are physical models or causal structures.

    \begin{definition}
        A \term{causal structure} is a directed acyclic graph (DAG) $\graph = \br{\nodes, \edges}$ with the extra distinction between two classes of nodes: the \term{latent nodes} $\nodes_L$ and \term{observed nodes} $\nodes_O$.
    \end{definition}
    The latent nodes of a causal structure correspond aim to represent completely general hidden variables that are hidden by some fundamental process or cannot be measured due to other limitations. The observed nodes represent random variables that are measurable and in practice will be observed as a collection of marginal contexts $\mscenario$. Any observation over this particular marginal scenario is expressible as a marginal model $\prob^{\mscenario}$. The field of causal inference aims to answer the looming question: \text{Is the observed marginal model compatible with the hypothesized causal structure?} \\

    In order to determine if a given marginal model $\prob^{\mscenario}$ is compatible with a causal structure $\graph$, a formal notion of \textit{compatibility} is warranted; first for a single marginal distribution $\prob[V]$ and then for a complete marginal model $\prob^{\mscenario} = \bc{\prob[V_1], \ldots, \prob[V_k]}$.

    \begin{definition}
        A set of \term{causal parameters} for a particular causal structure $\graph$ is the specification of a conditional distribution for every node $n \in \nodes$ given its parents in $\graph$.
        \[ \bc{\prob[n \mid \Pa[\graph]{n}] \mid n \in \nodes} \eq \label{eq:def:casual_param}\]
    \end{definition}

    \begin{definition}
        \label{def:compatible}
        A marginal distribution $\prob[V]$ is \term{compatible} with a causal structure $\graph$ if there exists a \textit{choice} of causal parameters such that $\prob[V]$ can be \textit{recovered} from the following series of operations:
        \begin{enumerate}
            \item Obtain a joint probability distribution over \textit{all} nodes of of the causal structure using the chosen casual parameters: $\prob[\nodes] = \prod_{n \in \nodes} \prob[n \mid \Pa[\graph]{n}]$.
            \item Marginalize over the latent nodes of $\graph$: $\prob[\nodes_O] = \sum_{\nodes_L} \prob[\nodes]$.
            \item Finally marginalize over the observed nodes not in $V$ to obtain $\prob[V]$: $\prob[V] = \sum_{\nodes_O \setminus V} \prob[\nodes_O]$
        \end{enumerate}
        If it is not possible for $\prob[V]$ to be recovered for \textit{any} choice of causal parameters, then $\prob[V]$ is \term{incompatible}. To generalize, a marginal model $\prob^{\mscenario}$ is compatible with $\graph$ if each distribution $\prob[V] \in \prob^{\mscenario}$ can be made compatible by the \textit{same} choice of causal parameters; otherwise it is incompatible.
    \end{definition}

    There are two reasons why a given marginal model $\prob^{\mscenario}$ would be deemed incompatible. One possibility is that the marginal model $\prob^{\mscenario}$ does not admit a joint distribution $\prob[\jointvar]$ such that it is impossible for requirement 3 of \cref{def:compatible} to hold\footnote{It is typical for $\nodes_O = \jointvar$. At the very least, it is demanded that $\jointvar \subseteq \nodes_O$.}. Alternatively, it could be possible that a joint distribution $\prob[\jointvar]$ exists but can not factor into causal parameters as per requirement 1 of \cref{def:compatible}. The former point-of-failure, namely the non-admission of a joint distribution, will be principle focus of this paper. Colloquially, the existence (or non-existence) of a joint distribution is referred to as \textit{the marginal problem}.

    \begin{definition}
        \label{def:marginal_problem}
        \term{The Marginal Problem:} Given a marginal model $\prob^{\mscenario} = \bc{\prob[V_1], \ldots, \prob[V_k]}$ marginal to the joint variables $\jointvar = \bigcup_{i} V_i$, does there exist a joint distribution $\prob[\jointvar]$ such that each given distribution $\prob[V_i]$ can be obtained by marginalizing $\prob[\jointvar]$?
        \[ \forall V \in \mscenario : \prob[V] = \sum_{\jointvar \setminus V} \prob[\jointvar] \eq \label{eq:def:marginal_problem}\]
        Following definition 2.3 in~\cite{Fritz_2011}, a marginal model is said to be \term{contextual} if it \textit{does not} admit a joint distribution $\prob[\jointvar]$ and \term{non-contextual} otherwise.
    \end{definition}

    There are two perspectives that can be taken when trying to answer the marginal problem. The first comes as stated by \cref{def:marginal_problem}; when inured with a particular marginal model $\prob^{\mscenario}$, can you construct a joint distribution that satisfies \cref{def:marginal_problem}? Alternatively and more abstractly, when provided a marginal \textit{mscenario} $\mscenario$ what is the set of all marginal models that are (non-)contextual? To facilitate future discussions, let $\s{NC}$ denote the set of marginal models that are non-contextual (admitting a joint distribution) and let $\s{C}$ denote the set of contextual marginal models. It becomes desirable to identifying conditions that constrain either $\s{NC}$ or $\s{C}$. We will continue to refer to a \term{complete solution marginal problem} as a finite set of necessary and sufficient conditions able to discern if a given marginal model $\prob^{\mscenario}$ is in $\s{NC}$ or $\s{C}$. These constraints typically take the form of linear inequalities over probabilities $\prob^{\mscenario}$.

    \begin{definition}
        A \term{$\s{NC}$-inequality} $I_{\mscenario}$ for a marginal scenario $\mscenario$ is a probabilistic inequality that is obeyed for every non-contextual marginal model $\prob^{\mscenario}$.
    \end{definition}

    In the \cref{sec:linear_program}, we will explore and briefly mention existing techniques for completely (or partially) solving the marginal problem. The basis of understanding being that \cref{eq:def:marginal_problem} is a \textit{linear} operation. In certain circumstances, \cref{eq:def:marginal_problem} can be represented as a matrix operation.

    \todo[TC]{Maybe define the possibilistic marginal problem for later} \\

    \subsection{Casting the Marginal Problem as a Linear Program}
    \label{sec:linear_program}

    Heretofore, our discussions of compatibility, contextuality and the marginal problem have be completely general, making no reference to the nature of the outcomes of the random variables $\jointvar$; whether they be discrete or continuous, finite or unbounded, or otherwise. \todo[TC]{It is entirely possible to obtain complete solutions to the marginal problem for continuous and unbounded measure spaces using entropic inequalities. [Elaborate and reference]} In this paper we restrict our attention to finite outcomes. In particular, we will focus on marginal scenarios that are finite but have a large number of outcomes. To assist with the subsequent analysis, we recall some familiar notions of probability theory together with new terminology and notation. \\

    \subsubsection{Events}
    To every discrete random variable $v$ there corresponds a prescribed set of \term{outcomes} $O_v$. We also define the set of all \term{events over $v$} $\Events{v}$\footnote{In the language of sheaf-theory, $\Events{v}$ is the \textit{sheaf of events}~\cite{Abramsky_2011}.} to be the set of all functions $s : \bc{v} \to O_v$ each representing the event that a measurement on $v$ was made where $s\br{v} \in O_v$ was observed.
    \[ \Events{v} \defined \bc{s : \bc{v} \to O_v} \label{eq:outcome_space} \]
    Evidently, $\Events{v}$ and $O_v$ have a one-to-one correspondence and this distinction can be confounding. There is rarely any harm in referring synonymously to either as outcomes. Nonetheless, a sheaf-theoretic treatment of contextuality~\cite{Abramsky_2011} demands the distinction.
    Specifically for this work, the distinction becomes essential for our discussion and exploit of marginal symmetries in \cref{sec:symmetry}. As a natural generalization we define the event over a collection of random variables $V = \bc{v_1, \ldots, v_n}$ in a parallel manner,
    \[ \Events{V} \defined \bc{s: V \to O_{V} \mid \forall i : s\br{v_i} \in O_{v_i} } \eq \label{eq:outcome_space}\]
    Each event $s$ can be compactly represented as a set of mappings over each element of $V$.
    \[ s = \bc{v_1 \mapsto s\br{v_1}, \ldots, v_k \mapsto s\br{v_k}} = \bc{v_i\mapsto s\br{v_i}}_{i = 1}^{k} \]
    The domain $\Dom{s}$ of an event $s$ is the set of random variables it valuates, i.e. if $s \in \Events{V}$, then $\Dom{s} = V$.
    \begin{example}
        Let $s \in \Events{V}$ be an event over $V$, a probability distribution $\prob[V]$ over $V$ will assign a non-negative probability $\prob[V][s]$ to $s$. Let $V = \bc{v_1, v_2}$ with each $v \in V$ having $3$ distinct outcomes $O_v = \bc{0,1,2}$. Furthermore let $f \in \Events{V}$ be the section $\br{v_1 \mapsto 2, v_2 \mapsto 1}$. Then the following notations are interchangeable,
        \[ \prob[][s] = \prob[\Dom{s}][s] = \prob[V][s] = \prob[v_1, v_2][s] = \prob[][v_1 \mapsto s_1\br{v_1}, v_2 \mapsto s_2\br{v_2}] = \prob[][v_1 \mapsto 2, v_2 \mapsto 1] = \prob[v_1v_2][21] \]
        Each of which should be read as ``the probability that $v_1$ got outcome $2$ and $v_2$ got outcome $1$''
    \end{example}
    \begin{definition}
    \label{def:section_restriction}
    For every $W \subseteq V$ and $s \in \Events{V}$, the \term{restriction of $s$ onto $W$} (denoted $s|_{W} \in \Events{W}$) is the event in $\Events{W}$ that \textit{agrees} with each of $s$'s assignments for variables in $W$.
    \[ \forall v \in W : s|_{W}\br{v} = s\br{v} \]
    \end{definition}
    \begin{definition}
    \label{def:section_compatible}
    The events $s_{V} \in \Events{V}$ and $s_{W} \in \Events{W}$ are said to be \term{compatible} events if they agree on their valuations of $V \cap W$. Compatibility will be denoted with the infix `$\com$'.
    \[ s_{V} \com s_{W} \iff {s_{V}}|_{V\cap W} = {s_{W}}|_{V\cap W} \]
    \end{definition}
    \begin{definition}
    \label{def:section_extendable}
    It $s_V \in \Events{V}$ and $s_W \in \Events{W}$ compatible events and $W \subseteq V$ then $s_W$ is said to be \term{extendable} to $s_V$. Extendability will be denoted with the infix `$\ext$'.
    \[ s_W \ext s_V \iff s_W \com s_V, W \subseteq V \]
    Extendability is dual to the notion of a \term{restriction} of $s_V$; $s_V$ can be restricted to $s_W$ ($s_V \res s_W$) if and only if $s_W$ is extendable to $s_V$.
    \end{definition}

    \begin{remark}
        The notation `$s \com t$' is used to indicate compatibility between $s \in \Events{V}$ and $t \in \Events{W}$ because of the following observation: $s$ and $t$ are compatible if and only if they are extensions of some event $r \in \Events{V \cap W}$.
        \[ s \com t \iff \exists r \in \Events{V \cap W} \mid s \res r \ext t \]
        In fact $r$ is uniquely defined.
        \[ \forall v \in V \cap W : r\br{v} = s\br{v} = t\br{v} \]
    \end{remark}

    For every marginal scenario $\mscenario = \bc{V_1, \ldots, V_k}$, it is useful to put special emphasis on the \term{joint events} $\Events{\jointvar}$. Similarly we define the \term{context events} $\bc{\Events{V} \mid V \in \mscenario}$. Finally, we elect to define the disjoint union of all context events are called the \term{marginal events} and by an abuse of notation we will denote this disjoint union as $\Events{\mscenario}$.
    \[ \Events{\mscenario} = \coprod_{V \in \mscenario} \Events{V} \]
    Each marginal section $m \in \Events{\mscenario}$ has a domain $\Dom{m} = V$ for some $V \in \mscenario$.
    \subsubsection{Linear Programs}
    \label{sec:linear_programs}

    We are now fully equipped to write the marginal problem as a matrix multiplication.
    \begin{definition}
        \label{def:incidence_matrix}
        The \term{incidence matrix} $M$ for a marginal scenario $\mscenario = \bc{V_1, \dots, V_k}$ is a bit-wise matrix where the columns are indexed by \textit{joint} events $j \in \Events{\jointvar}$ and the rows are events by \textit{marginal} sections $m \in \Events{\mscenario}$. The entries of $M$ are populated whenever a row index is extendable to a column index.
        \[ M_{m,j} = \de_{m \ext j} \defined \begin{cases}
            1 & m \ext j \\
            0 & \text{otherwise}
        \end{cases} \]
        Where $\de_{m \ext j}$ is termed the \term{extendability indicator}. The incidence matrix has $\abs{\Events{\jointvar}}$ columns, $\abs{\Events{\mscenario}} = \sum_{i} \abs{\Events{V_i}}$ rows and $k\abs{\Events{\jointvar}}$ non-zero entries.
    \end{definition}
    To illustrate this concretely, consider the following example:
    \begin{example}
        \label{ex:incidence_matrix}
        Let $\jointvar$ be $3$ binary variables $\jointvar = \bc{a,b,c}$ and $\mscenario$ be the marginal scenario $\mscenario = \bc{\bc{a,b}, \bc{b,c}, \bc{a,c}}$. The incidence matrix becomes:
        \[ M = \kbordermatrix{
            (a,b,c) \:\: \mapsto & (0,0,0) & (0,0,1) & (0,1,0) & (0,1,1) & (1,0,0) & (1,0,1) & (1,1,0) & (1,1,1) \\
            (a\mapsto0, b\mapsto0) & \kone & \kone & \kzer & \kzer & \kzer & \kzer & \kzer & \kzer \\
            (a\mapsto0, b\mapsto1) & \kzer & \kzer & \kone & \kone & \kzer & \kzer & \kzer & \kzer \\
            (a\mapsto1, b\mapsto0) & \kzer & \kzer & \kzer & \kzer & \kone & \kone & \kzer & \kzer \\
            (a\mapsto1, b\mapsto1) & \kzer & \kzer & \kzer & \kzer & \kzer & \kzer & \kone & \kone \\
            (b\mapsto0, c\mapsto0) & \kone & \kzer & \kzer & \kzer & \kone & \kzer & \kzer & \kzer \\
            (b\mapsto0, c\mapsto1) & \kzer & \kone & \kzer & \kzer & \kzer & \kone & \kzer & \kzer \\
            (b\mapsto1, c\mapsto0) & \kzer & \kzer & \kone & \kzer & \kzer & \kzer & \kone & \kzer \\
            (b\mapsto1, c\mapsto1) & \kzer & \kzer & \kzer & \kone & \kzer & \kzer & \kzer & \kone \\
            (a\mapsto0, c\mapsto0) & \kone & \kzer & \kone & \kzer & \kzer & \kzer & \kzer & \kzer \\
            (a\mapsto0, c\mapsto1) & \kzer & \kone & \kzer & \kone & \kzer & \kzer & \kzer & \kzer \\
            (a\mapsto1, c\mapsto0) & \kzer & \kzer & \kzer & \kzer & \kone & \kzer & \kone & \kzer \\
            (a\mapsto1, c\mapsto1) & \kzer & \kzer & \kzer & \kzer & \kzer & \kone & \kzer & \kone
        } \eq \label{eq:example_incidence}\]
    \end{example}

    \begin{definition}
        The \term{joint distribution vector} $\probvec^{\jointvar}$ for a probability distribution $\prob[\jointvar]$ is the vector whose entries are the positive, real-valued probabilities that $\prob[\jointvar]$ assigns to each joint event $j \in \Events{\jointvar}$. $\probvec^{\jointvar}$ shares the same indices as the \textit{column} indices of $M$.
        \[ \forall j \in \Events{\jointvar} : \probvec^{\jointvar}_j = \prob[\jointvar][j] \]
    \end{definition}
    \begin{definition}
        Analogously, \term{marginal distribution vector} $\probvec^{\mscenario}$ for a marginal model $\prob^{\mscenario} = \bc{P_{V_1}, \ldots, P_{V_k}}$ is the vector whose entries are probabilities over the set of marginal outcomes $\Events{\mscenario}$. $\probvec^{\mscenario}$ shares the same indices as the \textit{row} indices of $M$.
        \[ \forall m \in \Events{\mscenario} : \probvec^{\mscenario}_m = \prob[\Dom{m}][m] \]
    \end{definition}
    The marginal and joint distribution vectors are related via the incidence matrix $M$. Given a joint distribution vector $\probvec^{\jointvar}$ one can obtain the marginal distribution vector $\probvec^{\mscenario}$ by multiplying $M$ by $\probvec^{\jointvar}$.
    \[ \probvec^{\mscenario} = M \tcdot \probvec^{\jointvar} \eq \label{eq:incidence_matrix_use} \]
    Henceforth we will use $\probvec^{\jointvar}$ ($\probvec^{\mscenario}$) and $\prob[\jointvar]$ ($\prob_{\mscenario}$) interchangeably.

    \begin{remark}
        The particular ordering of the rows and columns of $M$ carries no importance, but it \textit{must} be consistent between $M$, $\probvec^{\jointvar}$ and $\probvec^{\mscenario}$.
    \end{remark}

    The marginal problem can now be rephrased in the language of the incidence matrix. Suppose one obtains a marginal distribution vector $\probvec^{\mscenario}$. The marginal problem becomes equivalent to the question: \textit{Does there exist a joint distribution vector $\probvec^{\jointvar}$ such that \cref{eq:incidence_matrix_use} holds?}

    \begin{definition}
        \label{def:marginal_linear_program}
        \term{The Marginal Linear Program} is the following linear program:
        \begin{alignat*}{2}
            & \text{minimize:} \quad&& \emptyset \tcdot x\\
            & \text{subject to:} && x \succeq 0 \\
            & && M \tcdot x = \probvec^{\mscenario}
        \end{alignat*}
        If this ``optimization''\footnote{``Optimization'' is presented in quotes here because the minimization objective is trivially always zero ($\emptyset$ denotes the null vector of all zero entries). The primal value of the linear program is of no interest, all that matters is its \textit{feasibility}.} is \textit{feasible}, then there exists a vector $x$ than can satisfy \cref{eq:incidence_matrix_use} and is a valid joint distribution vector. Therefore feasibility implies that $\probvec^{\jointvar} = x$. Moreover if the marginal linear program is \textit{infeasible}, then there \textit{does not} exist a joint distribution $\probvec^{\jointvar}$ over all random variables.
    \end{definition}
    \todo[TC]{Talk about about the linear solvers and how this is easy to implement.}
    \begin{definition}
        \label{def:dual_marginal_linear_program}
        \term{The Dual Marginal Linear Program} is the dual of \cref{def:marginal_linear_program} formulated via a procedure similar to~\cite{Lahaie_2008}:
        \begin{alignat*}{2}
            & \text{minimize:} \quad&& y \tcdot \probvec^{\mscenario}\\
            & \text{subject to:} && y \tcdot M \succeq 0
        \end{alignat*}
        Where $y$ is a real valued vector with the same length as $\probvec^{\mscenario}$.
    \end{definition}

    The dual marginal linear program not only answers the marginal problem for a specific marginal model $\probvec^{\mscenario}$ but as a by-product provides an inequality that witnesses its contextuality. If this is not obvious, first notice that the dual problem is \textit{never infeasible}; by choosing $y$ to be trivially the null vector $\emptyset$ of appropriate size, all constraints become satisfied. Secondly if the dual constraint $y \tcdot M \succeq 0$ holds \textit{and} the primal is feasible, then the following must hold:
    \[ y \tcdot \probvec^{\mscenario} =  y \tcdot M \tcdot x \geq 0 \eq \label{eq:dual_marginal_ineq} \]
    Therefore the \textit{sign} of the dual value $d \defined \min \br{y \tcdot \probvec^{\mscenario}}$ classifies a marginal model's contextuality. If $d < 0$ then \cref{eq:dual_marginal_ineq} is violated and therefore $\probvec^{\mscenario}$ is contextual. Likewise if $d \geq 0$ (satisfying \cref{eq:dual_marginal_ineq}), then $\probvec^{\mscenario}$ is non-contextual. \footnote{Actually, if $d \geq 0$ then it is exactly $d = 0$ due to the existence of the trivial $y = \emptyset$. This observation is an instance of the \textit{Complementary Slackness Property} of~\cite{Bradley_1977}. \comment[TC]{Is this really the CSP?} Moreover, if $d < 0$, then it is unbounded $d = -\inf$. This latter point becomes clear upon recognizing that for any $y$ with $d < 0$, another $y' = \al y$ can be constructed (with $\al > 1$) such that $d' = \al d < d$. Since a more negative $d'$ can always be found, it must be that $d$ is unbounded. This is a demonstration of the fundamental \textit{Unboundedness Property} of~\cite{Bradley_1977}; if the dual is unbounded, then the primal is infeasible. \comment[TC]{\href{http://web.stanford.edu/~ashishg/msande111/notes/chapter4.pdf}{Farkas's lemma} here?}}

    \begin{definition} An \term{infeasibility certificate}~\cite{Andersen_2001} is any vector $y$ that satisfies the constraint of \cref{def:dual_marginal_linear_program} and also permits violation of \cref{eq:dual_marginal_ineq} for some marginal distribution vector $\probvec^{\mscenario}$.
    \[ y \in \R^{\abs{\Events{\mscenario}}} : y \tcdot M \succeq 0, \quad y \tcdot \probvec^{\mscenario} < 0 \]
    Furthermore, for every $y$ satisfying $y \tcdot M \succeq 0$ there corresponds a \term{certificate inequality} that constraints the set of non-contextual marginal models $\s{NC}$,
    \[ \forall \probvec^{\mscenario} \in \s{NC} : y \tcdot \probvec^{\mscenario} \geq 0 \]
    \end{definition}

    \todo[TC]{Discuss Infeasibility Certificates basis} \\

    \todo[TC]{Discuss the certificate inequalities we found.} \\

    \todo[TC]{Discuss Compatibility, connection to cooperative games/resources, bell incompatibility?} \\

    \subsubsection{Fourier Motzkin Elimination \& Polytope Projection}

    The complete space of marginal models over $\mscenario$ can be partitioned into two spaces: contextual ($\s{C}$) and non-contextual marginal models ($\s{NC}$). The separation between these spaces is completely characterized by the system of inequalities associated with \cref{eq:incidence_matrix_use}. Explicitly,
    \begin{equation} \label{eq:halfspace_rep_incidence}
    \begin{alignedat}{2}
        \forall m \in \Events{\mscenario} :&&\quad \probvec^{\mscenario}_m - {\sum}_{j} M_{m, j} \probvec^{\jointvar}_j &\geq 0 \\
        \forall m \in \Events{\mscenario} :&&\quad -\probvec^{\mscenario}_m + {\sum}_{j} M_{m, j} \probvec^{\jointvar}_j &\geq 0 \\
        \forall j \in \Events{\jointvar} :&&\quad \probvec^{\jointvar}_j &\geq 0
    \end{alignedat}
    \end{equation}
    Geometrically \cref{eq:halfspace_rep_incidence} is a family of $2\abs{\Events{\mscenario}} + \abs{\Events{\jointvar}}$ half-planes bounding $\R^{\abs{\Events{\mscenario}}} \times \R^{\abs{\Events{\jointvar}}}$. However testing the contextuality of a marginal model $\probvec^{\mscenario}$ requires a collection of inequalities bounding just $\R^{\abs{\Events{\mscenario}}}$. Transforming \cref{eq:halfspace_rep_incidence} into the desired system of inequalities is the task of linear-quantifier elimination. A popular tool for linear quantifier elimination is Fourier-Motzkin elimination \cite{Dantzig_1973,Inflation,Abramsky_2012,jones_2004}. We will avoid expounding the Fourier-Motzkin procedure in this paper and instead reproduce a few of its notable features and consequences.

    \begin{definition} \cite[Section 12.2]{Schrijver_1998}
        The \term{Fourier-Motzkin elimination} procedure begins with a matrix $A$ and a vector $b$ defining a system of inequalities $S_1 = \bc{A\tcdot x \leq b}$ constraining a set of variables $x$ and symbolically eliminates a sub-set of theses variables $x \setminus y$. Upon completion, one obtains a new set of constraints $S_2 = \bc{A' y \leq b'}$ such that,
        \[ \exists y : A' y \leq b' \iff \exists x : A x \leq b \]
        Any solution $y$ of $S_2$ will permit at least one solution $x$ in $S_1$ (and vice versa).
    \end{definition}

    When applied to the marginal problem of \cref{eq:halfspace_rep_incidence}, the Fourier-Motzkin procedure eliminates all reference to the joint probabilities $\probvec^{\jointvar}_j$ and returns a system of inequalities over the marginal probabilities $\probvec^{\mscenario}_m$ representing a \textit{complete} solution to the marginal problem. Applying the Fourier-Motzkin procedure to completely solve the marginal problem is discussed in more detail in~\cite{Inflation} and~\cite{Fritz_2011}. \todo[TC]{Improve this discussion of FMP}. \\

    \todo[TC]{Discuss why FM is slow?}\\

    \begin{remark}
        In theory, Fourier-Motzkin offers a complete solution to the marginal problem for all marginal scenarios; its limitations are purely computational. In addition, the entire procedure only utilizes field operations \cite{Abramsky_2012} meaning that since \cref{eq:halfspace_rep_incidence} is a rational system of inequalities, then a complete solution to the marginal problem only requires a finite set of linear, rational coefficient inequalities.
    \end{remark}

    \begin{theorem}
        There exists a set of necessary and sufficient conditions

    \end{theorem}

    \todo[TC]{Logical Bell inequalities paper: only need to consider rational inequalities, actually integer inequalities, actually homogeneous inequalities} \\
    \todo[TC]{Discuss computational complexity} \\
    \todo[TC]{Mention what a complete solution looks like after using Fourier Motzkin}
    \subsubsection{Marginal Polytope}
    \begin{definition}
        Let $I_{1}$ and $I_{2}$ represent two
    \end{definition}
    \begin{definition}
        Let $\s{I}$
    \end{definition}

    \todo[TC]{Very briefly discuss polytope perspective and facet inequalities vs. non-facet inequalities}
    \todo[TC]{Define facets as irreducible inequalities}

    \todo[TC]{Maybe put the description of our task here. This will motivate the need for more scalable approaches.}

    \section{Transversal Inequalities}

    Section \ref{sec:linear_programs} discusses how one can obtain a $\s{NC}$-inequality that witnesses the contextuality of a particular marginal model $\probvec^{\mscenario}$ by writing the marginal problem as a linear program. Every $\probvec^{\mscenario} \in \s{NC}$ is witness-able by some certificate inequality. Albeit when a candidate marginal model $\probvec^{\mscenario}$ is not present, linear optimization offers no utility. \\

    This section aims to develop a procedure for finding \textit{all} certificate inequalities in a computationally efficient manner. Specifically by finding all certificate inequalities that are also facets of the marginal polytope. Before doing so, we will first summarize a method \todo[TC]{Method where this idea came from, original insight from Hardy?} that can be used to obtain a complete solution to the \textit{possibilistic} marginal problem~\cite{Mansfield_2012} and indirectly offer a \textit{partial} solution to the \textit{probabilistic} marginal problem.

    \subsection{Hardy Transversals}
    \label{sec:logical_inequalities}
    Let $a \in \Events{\mscenario}$ be any marginal event and $C = \bc{c_1, \ldots, c_{n}} \subseteq \Events{\mscenario}$ be a subset of the marginal events. We say that the subset $C$ forms a \term{logical extension} of $a$ if the following logical implication holds for \textit{all} marginal models $\probvec^{\mscenario}$.
    \[ a \implies c_1 \vee \ldots \vee c_n = \bigvee_{c \in C} c \eq \label{eq:1a_implication} \]
    Which can be dictated: \textit{Whenever the event $a$ occurs, at least one the events in $C$ occurs.} In this paper, the marginal event $a$ will be referred to as the \textit{antecedent} and $C$ as the \textit{consequent set}. To clarify, a marginal model $\probvec^{\mscenario}$ satisfies \cref{eq:1a_implication} if there always exists at least one $c \in C$ that is possible ($\probvec^{\mscenario}_c > 0$) whenever $a$ is possible. A marginal model $\probvec^{\mscenario}$ violates \cref{eq:1a_implication} whenever \textit{none} of the sections in $C$ are possible while $a$ is nonetheless still possible. Marginal models that violate logical statements such as \cref{eq:1a_implication} are known as \term{Hardy Paradoxes}~\cite{Inflation,Mansfield_2012,Mancinska_2014}. The concept of witnessing quantum incompatibility on a logical level has been analyzed thoroughly for decades~\cite{Greenberger_1990,Abramsky_2012}; being motivated by a greater sense of robustness compared to probabilistic constraints. \\

    All logical implications of the form \cref{eq:1a_implication} can be constructed by first selecting $a$ and then manufacturing a consequent set $C$ persuant to \cref{eq:1a_implication}.

    \begin{definition}
        \label{def:extendable_set}
        For each marginal event $m \in \Events{\mscenario}$ the \term{extendable set of $m$ into $\jointvar$} is the set of all joint events $j \in \Events{\jointvar}$ that are extensions of $m$.
        \[ \Ext[\jointvar]{m} = \bc{j \in \Events{\jointvar} \mid m \ext j} \]
    \end{definition}

    The extendable set of $m$ into $\jointvar$ is the subset of events in $\Events{\jointvar}$ that \textit{agree} with $m$ about valuations for variables in $\Dom{m}$. Practically speaking, the extendable set of $m$ into $\jointvar$ is represented by the $m$-th row of the incidence matrix $M$.

    \begin{lemma}
        \label{lemma:logical_extenable_set}
        Let $a \in \Events{\mscenario}$ and $\probvec^{\mscenario} \in \s{NC}$ be non-contextual. Then,
        \[ a \iff \bigvee_{j \in \Ext[\jointvar]{a}} j \eq \label{eq:lemma_logical_extendable_set}\]
    \end{lemma}
    The idea being if a joint distribution does exists, then the event $a$ represents partial knowledge of the entire system $\jointvar$; whenever $a$ is measured, exactly one joint event $j$ has occurred in reality and $j$ must be compatible with $a$. Likewise \cref{lemma:logical_extenable_set} applies to the consequent set $C \in \Events{\mscenario}$,
    \[ \bigvee_{c \in C} c \iff \bigvee_{c \in C} \bigvee_{j \in \Ext[\jointvar]{c}} j \eq \label{eq:logical_extension_C}\]
    If a subset $C$ can be found such that its extendable set \textit{covers} the extendable set of $a$,
    \[ \Ext[\jointvar]{a} \subseteq \bigcup_{c \in C} \Ext[\jointvar]{c} \eq \label{eq:logical_containment} \]
    Then the combination of \cref{eq:lemma_logical_extendable_set,eq:logical_extension_C,eq:logical_containment} yields \cref{eq:1a_implication}.  Computationally enumerating all logical extensions $C \subseteq \Events{\jointvar}$ for a particular antecedent $a \in \Events{\mscenario}$ corresponds iteratively selecting marginal sections $c \neq a$ for membership in $C$ until \cref{eq:logical_containment} is satisfies. This process is more generally known as a set-covering problem, or a hypergraph transversal problem~\cite{Inflation}, an idea we return to in the more general picture to follow in \cref{sec:dep}. \todo[TC]{Maybe use an example here to define a Hardy transversal.}\\

>>>>

\begin{document}
    \title{The Definite Extension Procedure for Large-Scale Marginal Satisfiability}
    \author{Thomas C. Fraser}
    \email{tcfraser@tcfraser.com}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada \\ University of Waterloo, Waterloo, Ontario, Canada}
    % \author{Elie Wolfe}
    % \email{ewolfe@perimeter@institute.ca}
    % \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada}
    \date{\today}
    \begin{abstract}
        \todo[TC]{Write an abstract}
    \end{abstract}
    \maketitle
    \tableofcontents

    \todo[TC]{Flesh out document from archive}
\end{document}
